---
title: "lax"
format: html
---
```{r}
library(tidyverse)
library(ggplot2)

WLAX <- read_csv("jtyrrell/lacrosse_women_ncaa_div1_2022_2023.csv") |> mutate(draw_pctg = draw_pctg * 100, 
                             win_pctg = win_pctg * 100)
```

# women regression model

```{r}
ggplot(WLAX, aes(x = draw_pctg, y = win_pctg)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  labs(title = "Winning the Draw vs. Winning the Game",
       x = "Draw Win Percentage",
       y = "Game Win Percentage") +
  theme_minimal()
```

# women regression line
```{r}
model <- lm(win_pctg ~ draw_pctg, data = WLAX)
summary(model)
```

-  for every 1% increase in draw win percentage, game win percentage increases by 2.4478 %

- p = 0.0000000000000002, depicting that draw percentage significantly affects win percentage

- R^2 = 0.5016, so the model explains 50.16% of the variation in win percentage

- Residual Standard Error: 
 RSE of 14.95 means that, on average, the model's predictions are off by about 14.95 units (percentage points)


# women polynomial regression model
```{r}
library(ggplot2)

ggplot(LAX, aes(x = draw_pctg, y = win_pctg)) +
  geom_point(color = "blue", alpha = 0.7) +  # Scatter plot
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "red", se = TRUE) +
  labs(title = "Polynomial Regression: Winning the Draw vs. Winning the Game",
       x = "Draw Win Percentage",
       y = "Game Win Percentage") +
  theme_minimal()

```

```{r}
model_poly <- lm(win_pctg ~ poly(draw_pctg, 2), data = LAX)
summary(model_poly)
```

# men regression model
```{r}
MLAX <- read_csv("jtyrrell/lacrosse_men_ncaa_div1_2022_2023.csv") |> mutate(faceoff_win_pct = faceoff_win_pct * 100,
                                                                          win_loss_pctg = win_loss_pctg * 100)

```

```{r}
ggplot(MLAX, aes(x = faceoff_win_pct, y = win_loss_pctg)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  labs(title = "Winning the Draw vs. Winning the Game",
       x = "Draw Win Percentage",
       y = "Game Win Percentage") +
  theme_minimal()
```
# men regression line
```{r}
model2 <- lm(win_loss_pctg ~ faceoff_win_pct, data = MLAX)
summary(model2)
```

-  for every 1% increase in draw win percentage, game win percentage increases by 1.5024 %

- p = 0.000001637, depicting that draw percentage significantly affects win percentage

- R^2 = 0.2815, so the model explains 28.15% of the variation in win percentage

- Residual Standard Error:
 RSE of 17.49 means that, on average, the model's predictions are off by about 17.49 units (percentage points)



- Men: for every 1% increase in draw win percentage, game win percentage increases by 1.5024%
- Women: for every 1% increase in draw win percentage, game win percentage increases by 2.4478%
- Winning the draw/faceoff has more impact on who wins the game for women compared to men


- Men: p = 0.000001637, depicting that draw percentage significantly affects win percentage
- Women: p = 0.00000000000000022, depicting that draw percentage significantly affects win percentage

- Men: R^2 = 0.2815, so the model explains 28.15% of the variation in win percentage
- Women: R^2 = 0.5016, so the model explains 50.16% of the variation in win percentage

- men:  RSE of 17.49 means that, on average, the model's predictions are off by about 0.8537 units (percentage points)
- women: RSE of 14.95 means that, on average, the model's predictions are off by about 0.709 units (percentage points)

Overall conclusions: 

- Winning the draw is a statistically significant predictor of game win percentage for both men and women. However, the effect is stronger for women, as indicated by a higher coefficient, a higher R² value, and a lower RSE.
- The model explains a larger portion of the variation in women's game win percentage than in men's, suggesting that draw win percentage is a more reliable predictor of game outcomes for women.


# both men and women

```{r}
WMLAX <- read_csv("jtyrrell/lacrosse_men_and_women_ncaa_div1_2022_2023.csv") |> mutate(faceoff_win_pct = faceoff_win_pct * 100,                                                              win_loss_pctg = win_loss_pctg * 100)

WMLAX$Division <- as.factor(WMLAX$Division)
```

# regression model with indicators
```{r}
ggplot(WMLAX, aes(x = faceoff_win_pct, y = win_loss_pctg, color = Division)) +
  geom_point(alpha = 0.7) +  
  geom_smooth(method = "lm", formula = y ~ poly(x, 1), se = TRUE) +  
  labs(title = "Regression Model: Draw Wins vs. Game Wins",
       x = "Draw Win Percentage",
       y = "Game Win Percentage") +
  theme_minimal()
```

```{r}
WMmodel <- lm(win_loss_pctg ~ faceoff_win_pct * Division, data = WMLAX)
summary(WMmodel)
```
# interaction model

```{r}
model_interaction <- lm(win_loss_pctg ~ faceoff_win_pct * Division, data = WMLAX)
summary(model_interaction)
```

- Men’s Effect (Baseline): Every 1% increase in faceoff win percentage increases men’s win percentage by 1.5024%.
Women’s Effect (Men’s Slope + Interaction Term): Every 1% increase in draw win percentage for women’s teams increases their win percentage by 1.5024% + 0.9454% = 2.4478%.
- This shows that for women, a 1% increase in draw win percentage leads to a 2.4478% increase in game win percentage...compared to 1.5024% for men
- both men’s and women’s teams benefit from winning more faceoffs/draws, but the impact is stronger for women
- p-values for the interaction and division terms are statistically significant (p < 0.01), confirming that gender plays a role in this relationship



